Subsegmental language detection in Celtic language text

### 2014-06-28 codeswitch-cltw-2014/codeswitch_camera_ready.tex
more and more and more
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55074 72bbbca6-d526-0410-a7d9-f06f51895060

\author{}
\author{Akshay Minocha \\
  IIIT Hyderabad  \\
  Hyderabad (India) \\
  {\small {\tt akshay.minocha@students.iiit.ac.in}} \\\And
  Francis M. Tyers \\
  Giellatekno / CLEAR \\
  UiT Norgga \'arktala\v{s} universitehta  \\
  9017 Romsa (Norway) \\
  {\small {\tt francis.tyers@uit.no}} \\}

@@

The work presented in this paper was motivated by the problems in normalising non-standard input
for the Celtic languages as a precursor to machine translation. When applying a normalisation 
strategy to a piece of text, it is necessary to first know the language of the piece of text you 
are applying it to.

There are several ways in which an informal text may contain parts in different languages.
Some of these are presented in Table~\ref{table:examples}.

@@

\begin{table}[h]
\begin{center}
\vspace{-0.3cm}
\begin{tabular}{ll}
\textbf{Code switching}: & You're a [Meirice\'{a}nach, c\'{e}n f\'{a}th] are you [foghlaim Gaeilge?!] \\
\textbf{Quotations}: & The anthem starts with the words [`Mae hen wlad fy nhadau\ldots'] \\
\textbf{Named entities}: & [Dr Jekyll] ha [Mr Hyde] embannet gant [\'{E}ditions Aber] \\
\textbf{Interjections}: & Hey, that's great, [diolch yn fawr!] \\
\textbf{Translations}: & Bloavezh mat d'an holl ! [Bonne ann\'{e} \`{a} tous !] \\
\end{tabular}
\caption{Some examples of text segments containing more than one language}
\label{table:examples}
\end{center}
\end{table}
\vspace{-0.65cm}

@@

In \newcite{chan2004detection} the authors have made use of the bi-phone probabilities and calculated them to measure a confidence metric, 
to \cite{lyu2006language} which has made use of a classification method, named syllable-based duration classification, 
which uses the tonal syllable properties along with the speech signals to help predict the code switch points. In \newcite{yeong2010language} 
 the authors use syllable structure information to identify words in code-switched text in Malay-English, however they did not 
recognise segments in running text, only identifying individual words.

@@

\begin{table}
\vspace{-0.65cm}
\begin{center}
\begin{tabular}{|l|r|r|r|r}
\hline
\multirow{2}{*}{\textbf{Pair}} & \multirow{2}{*}{\textbf{Language}}  & \multicolumn{2}{c|}{\textbf{Statistics} (\%)} \\\cline{3-4}
                &  &  Tokens & Segments \\
\hline
\multirow{2}{*}{Irish---English} & Irish & 332 & 40 \\
                                 & English & 379 & 42 \\
\hline
\multirow{2}{*}{Welsh---English} & Welsh & 419 & 64 \\
                                 & English & 378 & 66  \\
\hline
\multirow{2}{*}{Breton---French} & Breton & 388 & 54 \\
                                 & French & 379 & 53  \\
\hline
\end{tabular}
\end{center}
\caption{Document statistics of the annotated data used. }
\label{table:datastats}
\end{table}

@@

We use the character n-gram approach along with some heuristics which are relevant to our problem domain of identifying
segments for subsequent processing. We would like 
to both predict the code switched points but looking at the surrounding structure also decide the inclusion of them 
into the current or the next segment.\footnote{The software used in this experiment is available online at: \url{removed for review}.}

@@

This section describes the heuristics taken into account while performing the chunking of the input data according to 
the sequences of segments in different languages. 
In Figure ~\ref{fig:chunkal}, `chunks' represent the list of evaluated tuples of segments and their labelled language, `buffer' is the expanding segment.
 \textit{lang\_predict} corressponds to any function which is used to determine the language of the token. The flag 
variable helps in implementing the heuristic of minimum segment size while labelling chunks.
\begin{algorithm}
\caption*{\textsc{}}
%\caption{Chunking algorithm}\label{euclid}
{\fontsize{9}{9}\selectfont
\begin{algorithmic}[1]
%\Statex{}%Procedure{}{}
\State $\textit{buffer} = [\ ] \text{\ \ \  \#undecided expanding window of chunk}$
\State $\textit{chunks} = [\ ] \text{\ \ \  \#decided labelled segment}$
\State $\textit{buffer\_language} \gets \textit{lang\_predict}\text{(sentence[0])} \text{\ \ \ \#Language of first word}$
\State $\textit{flag} \gets \text{0}$
%\State $\text{for \ } \textit{word \ } \text{in \ } \textit{sentence:} $
\For{\texttt{word in sentence}}
\If {$ \textit{lang\_predict(word)} \text{==} \textit{buffer\_language} $}
        \If {$ \textit{flag} \text{=\ =\ 1} $}
        \State $\textit{buffer} += [\textit{word\_d,word}] $
        \State $\textit{flag} \gets \text{0}$
        \Else
        \State $\textit{buffer} += [\textit{word}] $
        \EndIf
        \EndIf
\If {$ \textit{lang\_prediction(word)} \text{!=} \textit{buffer\_langugae} $}
        \If {$ \textit{flag} \text{=\ =\ 0} $}
        \State $\textit{flag} \gets \text{1}$
        \State $\textit{word\_d} \gets \textit{word} $
        \State $\text{continue}$

@@

\EndIf
%\EndProcedure
%\Statex{}
\end{algorithmic}}
\end{algorithm}
\vspace{-0.6cm}
\begin{figure}[hb]
\def\@fs@post{}
\caption{Chunking Algorithm}
\label{fig:chunkal}
\end{figure}
%A flag is set at the beginning of the input text, and this moves forward to the position in the text upto where confident 
%chunking has been performed. For example, in case of a continuous prediction there would be no change and since the 
%predicted tag of the new word corresponds to that already of the expanding chunk, this word is also included in the 
%same chunk. In case of a doubt, the flag is set at the same position, and no final decision on labelling the word is 
%performed, to have more confidence on the code-switch, the process moves on to the next token, if it corresponds to 
%that of the previous chunk, then both these undecided words are labelled and included in the chunk, otherwise a change 
%is noted and a new chunk is created with the changed label. In this process, we are trying to identify the sub-segments 
%that we need, thus performing both tasks at once, the first being language prediction and second being the chunking decision.

\vspace{-0.8cm}
\subsection{Word-based prediction}
\label{wbased}

@@

\subsection{Word-based prediction with character backoff}
\label{wbasedbackoff}

A better way to predict the spans of the sub-segments of the text is to include the two methods of word and character-based techniques 
as described above. In case of the word being present in only one of the two monolingual word lists the classification is simple, 
but in case of a conflict, a character bigram backoff was introduced to help us disambiguate the language label. This method works well because the earlier heuristic approach of just labelling the word with the label of the span which is expanding, would mean less code-switch detection and more shift towards the majority class.

@@

\vspace{-0.2cm}

\section{Evaluation}
\label{sec:eval}

@@

\section{Results}
\vspace{-0.152cm}
\label{sec:results}
As described in Section~\ref{sec:method} the data collected from Twitter for the three language pairs, was processed using the techniques mentioned. The statistics of the same are given in Table~\ref{table:datastats}.

@@

\begin{table}
\begin{center}
\begin{tabular}{|l|r|r|r|}
\hline
\multirow{2}{*}{\textbf{System}} &  \multicolumn{3}{c|}{\textbf{Accuracy} (\%)} \\\cline{2-4}
       &   Irish---English & Welsh---English & Breton---French \\ 
\hline
\texttt{baseline} & 42.76 & 42.16 & 44.07 \\
\hline
\texttt{langid-3character} & 57.24 & 45.92 & 43.16 \\
\hline
\texttt{wordlist} & 79.75 & \textbf{74.28} & 83.96 \\
\hline
\texttt{character bigram} & 81.29 & 65.62 & 76.79 \\
\hline
\texttt{wordlist+character bigram} & \textbf{85.79} & 72.40 & \textbf{88.79} \\
\hline
\end{tabular}
\end{center}
\vspace{-0.3cm}
\caption{Accuracy of the systems over the three language pairs. The accuracy measures how often a token
  was assigned to the right language, independent of span.}
\label{table:accuracy}
\end{table}
In terms of per-word accuracy, the baseline performs as expected, given that half of the words in the test set are from a given language. 
The \texttt{langid.py} based methods allow for segmenting the text into chunks, but favour the majority language.\footnote{This could be because of the default models being used.} 
%While the precision and recall are low for the remaining models, we see that we are able to improve 
The performance by combining the wordlist based model with a character bigram model. And what is more, we are able to begin to not only 
identify particular words in a language, but also segments.

\section{Conclusions}
\label{sec:conclusions}

This paper has presented a very preliminary investigation into subsegment language identification in Celtic language texts. We have
proposed a model that chunks input text into segments and performs language identification on these segments at the same time. Precision and
recall are low, leaving a lot of room for further work. Some of our plans are as follows. We would like to annotate more test data,
at least 100 tweets in each language. Although In \cite{king2013labeling}, the authors label on a per word level, yet we would like to include supervised methods and features talked about in this research to improve our efficiency while dealing with segments. We would also like to attempt our method using higher order character n-gram models for backoff, and n-gram word language models for detection.

>>>>

\author{}
\author{Akshay Minocha \\
  IIIT Hyderabad  \\
  Hyderabad (India) \\
  {\small {\tt akshay.minocha@students.iiit.ac.in}} \\\And
  Francis M. Tyers \\
  Giellatekno \\ %/CLEAR \\
  UiT Norgga \'arktala\v{s} universitehta  \\
  9017 Romsa (Norway) \\
  {\small {\tt francis.tyers@uit.no}} \\}

@@

The work presented in this paper was motivated by the problems in normalising non-standard input
for the Celtic languages as a precursor to machine translation. When applying a normalisation 
strategy to a piece of text, it is necessary to first know the language of the piece of text you are applying it to.
%Table~\ref{table:examples} shows some examples of how code switching is present in text
%There are several ways in which an informal text may contain parts in different languages.
%Some of these are presented in Table~\ref{table:examples}.

@@

%\begin{table}[h]
%\begin{center}
%\vspace{-0.3cm}
%%\begin{tabular}{ll}
%\textbf{Code switching}: & You're a [Meirice\'{a}nach, c\'{e}n f\'{a}th] are you [foghlaim Gaeilge?!] \\
%\textbf{Quotations}: & The anthem starts with the words [`Mae hen wlad fy nhadau\ldots'] \\
%\textbf{Named entities}: & [Dr Jekyll] ha [Mr Hyde] embannet gant [\'{E}ditions Aber] \\
%\textbf{Interjections}: & Hey, that's great, [diolch yn fawr!] \\
%\textbf{Translations}: & Bloavezh mat d'an holl ! [Bonne ann\'{e} \`{a} tous !] \\
%\end{tabular}
%\caption{Some examples of text segments containing more than one language}
%\label{table:examples}
%\end{center}
%\end{table}
%\vspace{-0.65cm}

@@

In \newcite{chan2004detection} the authors have made use of the bi-phone probabilities and calculated them to measure a confidence metric, 
to \cite{lyu2006language} which has made use of named syllable-based duration classification, 
which uses the tonal syllable properties along with the speech signals to help predict the code switch points. In \newcite{yeong2010language} 
 the authors use syllable structure information to identify words in code-switched text in Malay-English, however they did not 
recognise segments in running text, only identifying individual words.

@@

\begin{table}
\begin{center}
\begin{tabular}{|l|r|r|r|r}
\hline
\multirow{2}{*}{\textbf{Pair}} & \multirow{2}{*}{\textbf{Language}}  & \multicolumn{2}{c|}{\textbf{Statistics} (\%)} \\\cline{3-4}
                &  &  Tokens & Segments \\
\hline
\multirow{2}{*}{Irish---English} & Irish & 332 & 40 \\
                                 & English & 379 & 42 \\
\hline
\multirow{2}{*}{Welsh---English} & Welsh & 419 & 64 \\
                                 & English & 378 & 66  \\
\hline
\multirow{2}{*}{Breton---French} & Breton & 388 & 54 \\
                                 & French & 379 & 53  \\
\hline
\end{tabular}
\end{center}
\caption{Document statistics of the annotated data used. }
\label{table:datastats}
\end{table}

@@

We use the character n-gram approach along with some heuristics which are relevant to our problem domain of identifying
segments for subsequent processing. We would like 
to both predict the code switched points but looking at the surrounding structure also decide the inclusion of them 
into the current or the next segment.
%\footnote{The software used in this experiment is available online at: \url{removed for review}.}

@@

This section describes the way we apply heuristics to segment and label the input text.
%the heuristics taken into account while performing the chunking of the input data according to 
%the sequences of segments in different languages. 
In Figure ~\ref{fig:chunkal}, `chunks' represent the list of evaluated tuples of segments and their labelled language, `buffer' is the expanding segment.
 \textit{lang\_predict} corressponds to any function which is used to determine the language of the token. The flag 
variable helps in implementing the heuristic of minimum segment size while labelling chunks.
\begin{algorithm}
\caption*{\textsc{}}
%\caption{Chunking algorithm}\label{euclid}
{\fontsize{9}{9}\selectfont
\begin{algorithmic}[1]
%\Statex{}%Procedure{}{}
\State $\textit{buffer} = [\ ] \text{\ \ \  \#undecided expanding window of chunk}$
\State $\textit{chunks} = [\ ] \text{\ \ \  \#decided labelled segment}$
\State $\textit{buffer\_language} \gets \textit{lang\_predict}\text{(sentence[0])} \text{\ \ \ \#Language of first word}$
\State $\textit{flag} \gets \text{0}$
%\State $\text{for \ } \textit{word \ } \text{in \ } \textit{sentence:} $
\For{\texttt{word in sentence}}
\If {$ \textit{lang\_predict(word)} \text{==} \textit{buffer\_language} $}
        \If {$ \textit{flag} \text{=\ =\ 1} $}
        \State $\textit{buffer} += [\textit{word\_d,word}] $
        \State $\textit{flag} \gets \text{0}$
        \Else
        \State $\textit{buffer} += [\textit{word}] $
        \EndIf
        \EndIf
\If {$ \textit{lang\_prediction(word)} \text{!=} \textit{buffer\_langugae} $}
        \If {$ \textit{flag} \text{=\ =\ 0} $}
        \State $\textit{flag} \gets \text{1}$
        \State $\textit{word\_d} \gets \textit{word} $
        \State $\text{continue}$

@@

\EndIf
%\EndProcedure
%\Statex{}
\end{algorithmic}}
\end{algorithm}
\begin{figure}[hb]
\def\@fs@post{}
\caption{Chunking Algorithm}
\label{fig:chunkal}
\end{figure}
%A flag is set at the beginning of the input text, and this moves forward to the position in the text upto where confident 
%chunking has been performed. For example, in case of a continuous prediction there would be no change and since the 
%predicted tag of the new word corresponds to that already of the expanding chunk, this word is also included in the 
%same chunk. In case of a doubt, the flag is set at the same position, and no final decision on labelling the word is 
%performed, to have more confidence on the code-switch, the process moves on to the next token, if it corresponds to 
%that of the previous chunk, then both these undecided words are labelled and included in the chunk, otherwise a change 
%is noted and a new chunk is created with the changed label. In this process, we are trying to identify the sub-segments 
%that we need, thus performing both tasks at once, the first being language prediction and second being the chunking decision.

\vspace{-0.7cm}
\subsection{Word-based prediction}
\vspace{-0.132cm}
\label{wbased}

@@

\subsection{Word-based prediction with character backoff}
\vspace{-0.152cm}
\label{wbasedbackoff}

%A better way to predict the spans of the sub-segments of the text is to include the two methods of word and character-based techniques 
%as described above.
In case of the word being present in only one of the two monolingual word lists the classification is simple, 
but in case of a conflict, a character bigram backoff was introduced to help us disambiguate the language label. This method works well because the earlier heuristic approach of just labelling the word with the label of the span which is expanding, would mean less code-switch detection and more shift towards the majority class.

@@

\section{Evaluation}
\vspace{-0.152cm}
\label{sec:eval}

@@

\section{Results}
\label{sec:results}
As described in Section~\ref{sec:method} the data collected from Twitter for the three language pairs, was processed using the techniques mentioned. The statistics of the same are given in Table~\ref{table:datastats}.

@@

\begin{table}
\begin{center}
\begin{tabular}{|l|r|r|r|}
\hline
\multirow{2}{*}{\textbf{System}} &  \multicolumn{3}{c|}{\textbf{Accuracy} (\%)} \\\cline{2-4}
       &   Irish---English & Welsh---English & Breton---French \\ 
\hline
\texttt{baseline} & 42.76 & 42.16 & 44.07 \\
\hline
\texttt{langid-3character} & 57.24 & 45.92 & 43.16 \\
\hline
\texttt{wordlist} & 79.75 & \textbf{74.28} & 83.96 \\
\hline
\texttt{character bigram} & 81.29 & 65.62 & 76.79 \\
\hline
\texttt{wordlist+character bigram} & \textbf{85.79} & 72.40 & \textbf{88.79} \\
\hline
\end{tabular}
\end{center}
\vspace{-0.3cm}
\caption{Accuracy of the systems over the three language pairs. The accuracy measures how often a token
  was assigned to the right language, independent of span.}
\label{table:accuracy}
\end{table}
In terms of per-word accuracy, the baseline performs as expected, given that half of the words in the test set are from a given language. 
The \texttt{langid.py} based methods allow for segmenting the text into chunks, but favour the majority language.
%\footnote{This could be because of the default models being used.} 
%While the precision and recall are low for the remaining models, we see that we are able to improve 
The performance by combining the wordlist based model with a character bigram model. And what is more, we are able to begin to not only 
identify particular words in a language, but also segments.

\section{Conclusions}
\vspace{-0.152cm}
\label{sec:conclusions}

This paper has presented a very preliminary investigation into subsegment language identification in Celtic language texts. We have
proposed a model that chunks input text into segments and performs language identification on these segments at the same time. Precision and
recall are low, leaving a lot of room for further work.
% Some of our plans are as follows. We would like to annotate more test data,
%at least 100 tweets in each language.
 Although In \cite{king2013labeling}, the authors label on a per word level, yet we would like to include supervised methods and features talked about in this research to improve our efficiency while dealing with segments. We would also like to attempt our method using higher order character n-gram models for backoff, and n-gram word language models for detection and on more annotated data.
